---
title: "GEO_tabeller"
author: "Vegard"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/helseprofil/backend/main/geomaster/R/functions_GeoMaster.R")

# Database file paths
root <- "O:/Prosjekt/FHP/PRODUKSJON/STYRING/"
khelsa <- "KHELSA.mdb"
geokoder <- "raw-khelse/geo-koder.accdb"
```

```{r}
con <- khfunctions:::connect_khelsa()
KnrHarmOld <- RODBC::sqlQuery(con, "SELECT * FROM KnrHarm")
GeoKoderOld <- RODBC::sqlQuery(con, "SELECT * FROM GeoKoder")
```


# Update GEO-tables for KHELSA

The following codes are used to generate updated versions of the tables `GeoKoder` and `KnrHarm` in `KHELSA` (KHELSA.mdb)

## Update KnrHarm table

- Connects to the databases and read relevant tables
- Update recoding table according to selected year
- Perform quality control to see if
    - All existing rows are kept or correctly updated
    - There are any missing values in the `GEO` or `GEO_omk` columns
    - All values in `GEO_omk` are valid according to `tblGeo` for the selected year
    
- Set `write = TRUE` to overwrite the table in Access. 
    - This will replace the original table, so first make sure that the table is correct by running the code with `write = FALSE`

```{r}
KnrHarmNew <- KnrHarmUpdate(year = 2025,
                            basepath = root,
                            khelsapath = khelsa,
                            geokoderpath = geokoder,
                            write = FALSE)
```

## Update GeoKoder

```{r}
GeoKoderNew <- GeoKoderUpdate(year = 2025,
                              basepath = root,
                              khelsapath = khelsa,
                              geokoderpath = geokoder,
                              write = TRUE)
```
