---
title: "Levekårssoner"
output: html_document
date: "2024-09-05"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
#options(scipen = 2)
options(digits = 2)
options(scipen=999)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
pacman::p_load(tidyverse, scales, Hmisc, stringr, jsonlite)

source("C:/Users/HANN/OneDrive - Helsedirektoratet/R_local/00_scripts/local_functions.R")
```


# Data


## Levekårssoner

Geokodes inkluding names, norwegian and samic


```{r}
geo2026 <- read_delim("C:/Users/HANN/OneDrive - Helsedirektoratet/R_local/00_data/GeoKoder_tblGeo_2026.csv", delim = ";", locale = locale(encoding = 'latin1'))

```
  


# Dataprep 

Create file with following structure as input for create-json script

LKS under kommuner are level 4, LKS under bydeler are level 5

geolev1_code  geolev1_name  geolev2_code  geolev2_name  geolev3_code  geolev3_name  geolev4_code  geolev4_name  geolev5_code  geolev5_name
0	Hele landet	03	Oslo (fylke)	0301	Oslo	030101	Gamle Oslo	0301010001	Oslo LKS 010001
0	Hele landet	03	Oslo (fylke)	0301	Oslo	030101	Gamle Oslo	0301010002	Oslo LKS 010002
0	Hele landet	03	Oslo (fylke)	0301	Oslo	030101	Gamle Oslo	0301010003	Oslo LKS 010003


## Cleaning


Remove kommuner/bydeler with ONE LKS

```{r}
removeLKS_code <- geo2026 %>% 
  filter(!is.na(levekaar)) %>% 
  separate(code, sep = 6, into = "geogroup", remove = FALSE) %>% 
  group_by(geogroup) %>% 
  filter(n() == 1) %>% 
  pull(code)
```


remove "uoppgitt", "99" codes

```{r}
geo2026 %>% 
  filter(name == "Uoppgitt")

geo2026 %>% 
  filter(grepl("99", code))
```

```{r}
geo2026_clean <- geo2026 %>%   
  filter(!grepl("99", code),
         !code %in% c(removeLKS_code)) %>% 
  select(-c(grunnkrets, validTo))
```



## Kommuner with LKS

Only 66 kommuner with LKS
Only 58 approved kommuner with LKS (removed the ones with one LKS)

```{r}
kommuner_lks <- geo2026_clean %>% 
  filter(!is.na(levekaar)) %>% 
  pull(kommune) %>% 
  unique()

kommuner_lks
length(kommuner_lks)
```

## Bykommuner

```{r}
bykommuner <- geo2026_clean %>% 
  filter(!is.na(bydel)) %>% 
  pull(kommune) %>% 
  unique()

bykommuner_lks <- bykommuner[bykommuner %in% kommuner_lks]

```

```{r}
version

```






# JSON input nested GEO table

- add geo 0, Hele landet
- add fylke-, kommune and bydel name
- add "(fylke)" for Oslo, due to duplicate error when inporting json-file to statbank


## Geo names

Duplicate kommuner Herøy and Våler allready contains fylkename from SSB

```{r}

geo2026_clean %>% 
  filter(grepl("Herøy|Våler", name))
```

```{r}
fylke <- geo2026_clean %>% 
  filter(level == "fylke") %>% 
  select(code, name) %>% 
  rename(fylke_name = name,
         fylke = code) %>% 
  mutate(fylke_name = ifelse(fylke == "03", paste0(fylke_name, " (fylke)"), fylke_name))

kommune <- geo2026_clean %>% 
  filter(level == "kommune") %>% 
  select(code, name) %>% 
  rename(kommune_name = name,
         kommune = code) 

bydel <- geo2026_clean %>% 
  filter(level == "bydel") %>% 
  select(code, name) %>% 
  rename(bydel_name = name,
         bydel = code)

```

```{r}
fylke
kommune
bydel
```


```{r}

geo2026_clean_names <- geo2026_clean %>% 
    mutate(geolev1_code = "0",
         geolev1_name = "Hele landet") %>% 
  left_join(fylke) %>% 
  left_join(kommune) %>% 
  left_join(bydel)

```


## Bykommuner - LKS at level 5

```{r}
jsoninput_bykommuner <- geo2026_clean_names %>% 
  filter(kommune %in% bykommuner,
         level == ifelse(kommune %in% bykommuner_lks, "levekaar", "bydel"))  %>% 
  mutate(geolev5_name = ifelse(kommune %in% bykommuner_lks, name, NA_character_)) %>% 
  rename(geolev2_code = fylke,
         geolev2_name = fylke_name,
         geolev3_code = kommune,
         geolev3_name = kommune_name,
         geolev4_code = bydel,
         geolev4_name = bydel_name,
         geolev5_code = levekaar) %>% 
  select(geolev1_code, geolev1_name, geolev2_code, geolev2_name, geolev3_code, geolev3_name, geolev4_code, geolev4_name, geolev5_code, geolev5_name)
  
```


## Kommuner - LKS at level 4

```{r}
jsoninput_kommuner <- geo2026_clean_names %>% 
  filter(!kommune %in% bykommuner,
         level == ifelse(kommune %in% kommuner_lks, "levekaar", "kommune"))  %>% 
  mutate(geolev4_name = ifelse(kommune %in% kommuner_lks, name, NA_character_)) %>% 
  rename(geolev2_code = fylke,
         geolev2_name = fylke_name,
         geolev3_code = kommune,
         geolev3_name = kommune_name,
         geolev4_code = levekaar) %>%
  mutate(geolev5_code = NA_character_,
         geolev5_name = NA_character_) %>% 
  select(geolev1_code, geolev1_name, geolev2_code, geolev2_name, geolev3_code, geolev3_name, geolev4_code, geolev4_name, geolev5_code, geolev5_name)
```


```{r}
jsoninput <- rbind(jsoninput_bykommuner, jsoninput_kommuner) %>% 
  arrange(geolev2_code, geolev3_code, geolev4_code)


jsoninput %>% 
  summarise(n = n_distinct(geolev3_code))

  
```




# Convert to Json format

## make json file

```{r}
data <- jsoninput
templatename <- "GEO LKS, 2026"
filename <- "LKS_KB_GEO2026.json"

# Load required libraries
library(tidyverse)
library(jsonlite)

# Recursive function to build the hierarchical JSON structure
build_hierarchy <- function(data, level = 1) {
  level_code <- paste0("geolev", level, "_code")
  level_name <- paste0("geolev", level, "_name")
  
  next_level <- level + 1
  
  # Get unique levels for the current geolev
  unique_levels <- data %>%
    distinct(!!sym(level_code), !!sym(level_name)) %>%
    filter(!is.na(!!sym(level_code)), !is.na(!!sym(level_name)))  # Exclude rows with NA code or name
  
  # Build the hierarchy recursively
  categories <- unique_levels %>%
    pmap(function(...) {
      row_data <- tibble(...)
      children_data <- data %>%
        filter(!!sym(level_code) == row_data[[1]])
      
      # Recursively build the children, stop at geolev5
      if (next_level <= 5) {
        children <- build_hierarchy(children_data, next_level)
      } else {
        children <- list()
      }
      
      # Build the current category node
      category <- list(
        Label = row_data[[2]],
        Value = row_data[[1]]
      )
      
      # Only include Children if there are valid children
      if (length(children) > 0) {
        category$Children <- children
      }
      
      return(category)
    })
  
  return(categories)
}

# Build the hierarchy starting from geolev1
hierarchy <- build_hierarchy(data)

# Final JSON structure
output <- list(
  TemplateName = templatename,
  Dimension = list(
    Code = "GEO",
    Label = "Geografi",
    Categories = hierarchy
  )
)

# Convert to JSON format
json_output <- toJSON(output, pretty = TRUE, auto_unbox = TRUE)

# Write to a JSON file
write(json_output, filename)

# Print the JSON output
cat(json_output)
```

